---
title: "EucFACE P Synthesis"
author: "Mingkai Jiang"
date: "1/29/2018"
output:
  html_document: default
---


# 1. Description

This is the progress report of the EucFACE phosphorus (P) synthesis project. All results presented in this report is automatically linked with the code and the data. 

This file attempts to summarize P-related pools, fluxes, concentrations, their treatment and inter-ring variability, and their temporal patterns. 

The objective of this project is to have a proper account of the P balance for the EucFACE experiment, and to investigate the P by CO2 interactions over time (if enough data is available).

All the data used in this study are available at HIEv (at least that's the goal). 

Here I am summarizing the time by treatment effect, for each individual variable. I will summarize the ring-specific result separately. 


# 2. Methods
### 2.1 Procedure
This procedure is a snapshot of the work. It might be easier just to follow the codes. 

    - Create a list of variables for the P synthesis.
    - For each variable, only consider ring, date, and sometimes depth as independent factors.
    - Calculate P concentrations first.
    - With Carbon balance output, calculate P pools and fluxes.
    - Calculate P budgeting variables, e.g. total standing P stock, P retranslocation coefficient, etc.
    - Make summary tables.
    - Interpret data, e.g. time-series trend, treatment effect, extreme response point, etc. 

### 2.2 Variables
Here, I am summarizing all the P-related pool, flux and concentration variables. 
Currently, this list is in-comprehensive. Will update continuously as the project proceeds. 

    - Soil P concentration and pool (different chemical forms and accessibilities)
    - Microbial P concentration and pool
    - Mycorrhizal P concentration and pool
    - Canopy P concentration, production and pool
    - Wood P concentration, production and pool
    - Fineroot P concentration, production and pool
    - Litterfall P concentration and production
    - P mineralization rate
    - Frass P concentration and production
    - Understorey P concentration, production and pool
    - Standing P stock
    - P requirement
    - Total P retranslocation 
    - Leaf P retranslocation coefficient
    - P uptake
    - P uptake / requirement
    - MRT
    - Standing PUE
    
### 2.3 How to derive P pools and fluxes based on concentration

The challenges with the P synthesis project is that, the P concentration measurements are sporadic and infrequent, we cannot simply taking the averages of concentration measurements and multiply it with the biomass data to obtain a P flux. 

For pool variables, i.e. wood P pool, we may be able to get away with using just one concentration measurement, given that the concentration of this pool is rather static and we only have one measurement in time avaiable after all. Taking this into consideration, we need to firstly check if the P concentration for a particular vegetation/soil component is static or not over time. If the P concentration is most likely to be static, then we are fine with using average values to extrapolate. On the other hand, for pools that have rather fluctuating P concentration over time, we probably want to establish some sort of relationship with meteorological data for extrapolation purpose, or try to find the biomass peak and the P concentration associated with this time point, assuming that the product of biomass peak and P concentration returns the highest P pool within this year.  

For fluxes, we also need to see if the concentration is variable over time or not. For those that are more fluctuating with good data coverage, e.g. frass P concentration, we can simply calculate a monthly/quarterly P concentration value and use this value to exrtapolate with C production for its corresponding P fluxes. For those that are fluctuating but with relatively less data coverage and with infrequent measurement frequency, e.g. leaflitter fall P concentration (4 time points over 3 years), we will have to assume the one time point is reflecting the annual P concentration, or otherwise we wouldn't be able to proceed. 

The other way to treat the complexity of variable P concentration over time is make a **sensitivity test**. We can use average P concentration, minimum and maximum P concentrations to respectively calculate the P pools and fluxes. This way we are probably more certain with regard to the range of uncertainties. 

### 2.4 Treatment of canopy biomass
Canopy biomass varies over time. At the same time, P concentration in canopy also varies. 



# 3. P concentration calculations. 
``` {r, echo=FALSE, message=FALSE, warning=FALSE}
    ### source the functions to start the calculations.
    source("programs/prepare.R")
    
    ### assign aco2 and eco2 rings
    aCO2_rings <- c(2,3,6)
    eCO2_rings <- c(1,4,5)

```

### 3.1 Canopy P concentration
Canopy P concentration summarized by ring over the experimental period:
``` {r, echo=FALSE}
    ### calculate the variable
    canopy_p_concentration <- make_canopy_p_concentration()
    
    ### make the ring over date plot
    p <- ggplot(canopy_p_concentration,
                aes(Date, PercP, color=factor(Ring))) +   
            geom_point(size = 5) + geom_line() + 
            xlab("Date") + ylab("Canopy P Concentration (%)") + 
            scale_color_manual(values=c("#FF7F50", "#00FFFF", "#6495ED",
                                        "#FF4040", "#8B0000", "#0000FF"))
    plot(p)

```

Canopy P concentration summarized by CO2 treatment over the experimental period:
``` {r, echo=FALSE}
    ### assign CO2 treatment labels
    for (i in aCO2_rings) {
        canopy_p_concentration[canopy_p_concentration$Ring == i, "CO2"] <- "aCO2"
    }
    for (i in eCO2_rings) {
        canopy_p_concentration[canopy_p_concentration$Ring == i, "CO2"] <- "eCO2"
    }
    
    ### make the treatment over date plot
    p <- ggplot(canopy_p_concentration,
                aes(Date, PercP, color=factor(CO2))) +   
            geom_point(size = 5) + 
            xlab("Date") + ylab("Canopy P Concentration (%)")
    plot(p)

```

Visually, there is **no** CO2 treatment effect. 



### 3.2 Wood P concentration
Wood P concentration summarized by treatment, and we only have one date:

``` {r, echo=FALSE}
    ### calculate the variable
    wood_p_concentration <- make_wood_p_concentration()
    
    ### assign CO2 treatment labels
    for (i in aCO2_rings) {
        wood_p_concentration[wood_p_concentration$Ring == i, "CO2"] <- "aCO2"
    }
    for (i in eCO2_rings) {
        wood_p_concentration[wood_p_concentration$Ring == i, "CO2"] <- "eCO2"
    }
    
    ### make the ring over date plot
    p <- ggplot(wood_p_concentration,
                aes(CO2, PercP)) +   
            geom_point(size = 5) +
            xlab("Treatment") + ylab("Wood P Concentration (%)")
    plot(p)

```

Visually, we can't distinguish the difference between CO2 treatment.



### 3.3 Fine root P concentration

There are multiple depths (i.e. 0-10cm, 10-30cm), currently we are averaging the concentrations over depths. 

Fineroot P concentration summarized by treatment over time:

``` {r, echo=FALSE}
    ### calculate the variable
    fineroot_p_concentration <- make_fineroot_p_concentration()
    
    ### assign CO2 treatment labels
    for (i in aCO2_rings) {
        fineroot_p_concentration[fineroot_p_concentration$Ring == i, "CO2"] <- "aCO2"
    }
    for (i in eCO2_rings) {
        fineroot_p_concentration[fineroot_p_concentration$Ring == i, "CO2"] <- "eCO2"
    }
    
    ### make the treatment over date plot
    p <- ggplot(fineroot_p_concentration,
                aes(Date, PercP, color=factor(CO2))) +   
            geom_point(size = 5) + 
            xlab("Date") + ylab("Fineroot P Concentration (%)")
    plot(p)

```


### 3.4 Leaflitter P concentration
Measurements are taken for both scenesced leaves and leaflitter. 
Scenesced leaves are defined as those that are still (barely) attached to the tree, so there is a tree ID associated with each measurement. 
Leaflitter are those collected from the litter baskets.

Note, these samples were not taken on the same date; nor were they taken on the same dates as the fresh green leaves. 

We firstly need to distinguish the difference between the two. 

``` {r, echo=FALSE, warning=FALSE}
    df <- read.csv("download/FACE_P0020_RA_leafP-Eter_20130201-20151115_L1.csv")
    
    ### setting up the date
    df$Date <- paste0("1-", as.character(df$Campaign))
    df$Date <- as.Date(df$Date, "%d-%b-%y")
    df.litter <- subset(df, Type == "Leaf litter")
    df.dead <- subset(df, Type == "sceneced leaf")
    myDF <- rbind(df.litter, df.dead)

    ### plot
    p <- ggplot(myDF,
                aes(Date, PercP, color=factor(Type))) +   
            geom_point(size = 5) +
            xlab("Date") + ylab("Leaflitter P Concentration (%)")
    plot(p)
```

Next, both leaflitter and senesced leaves were grouped together as leaflitter. 
The CO2 by date plot is:
``` {r, echo=FALSE}
    ### calculate the variable
    leaflitter_p_concentration <- make_leaflitter_p_concentration()
    
    ### assign CO2 treatment labels
    for (i in aCO2_rings) {
        leaflitter_p_concentration[leaflitter_p_concentration$Ring == i, "CO2"] <- "aCO2"
    }
    for (i in eCO2_rings) {
        leaflitter_p_concentration[leaflitter_p_concentration$Ring == i, "CO2"] <- "eCO2"
    }
    
    ### make the treatment over date plot
    p <- ggplot(leaflitter_p_concentration,
                aes(Date, PercP, color=factor(CO2))) +   
            geom_point(size = 5) + 
            xlab("Date") + ylab("Leaflitter P Concentration (%)")
    plot(p)

```


### 3.5 Understorey P concentration
There are two understorey species: Cymbopogon refractus and Microlaena stipoides. We need to differentiate P concentration difference between the two, as we don't have their relative abundance information. 

``` {r, echo=FALSE} 
    ### read in the file
    df <- read.csv("download/FACE_P0019_RA_leafP-understory_20130509-20151030_L1.csv")
    
    ### Assign date format
    df$Date <- as.Date(df$Date, "%d-%b-%y")
    
    ### Plot species comparisons
    p <- ggplot(df,
            aes(Date, PercP, color=factor(Species))) +   
        geom_point(size = 5) +
        xlab("Date") + ylab("Understorey P Concentration (%)")
    plot(p)
```

Now, we can group both species together and plot understorey P concentration over time.
``` {r, echo=FALSE} 
    understorey_p_concentration <- make_understorey_p_concentration()

    ### assign CO2 treatment labels
    for (i in aCO2_rings) {
        understorey_p_concentration[understorey_p_concentration$Ring == i, "CO2"] <- "aCO2"
    }
    for (i in eCO2_rings) {
        understorey_p_concentration[understorey_p_concentration$Ring == i, "CO2"] <- "eCO2"
    }

    ### make the treatment over date plot
    p <- ggplot(understorey_p_concentration,
                aes(Date, PercP, color=factor(CO2))) +   
            geom_point(size = 5) +
            xlab("Date") + ylab("Understorey P Concentration (%)")
    plot(p)

```

### 3.6 Frass P concentration
Frass production 

``` {r, echo=FALSE, message=FALSE, warning=FALSE}
    ### calculate the variable
    frass_p_concentration <- make_frass_p_concentration()
    
    ### assign CO2 treatment labels
    for (i in aCO2_rings) {
        frass_p_concentration[frass_p_concentration$Ring == i, "CO2"] <- "aCO2"
    }
    for (i in eCO2_rings) {
        frass_p_concentration[frass_p_concentration$Ring == i, "CO2"] <- "eCO2"
    }
    
    ### make the ring over date plot
    p <- ggplot(frass_p_concentration,
                aes(Date, PercP, color=factor(CO2))) +   
            geom_point(size = 5) +
            xlab("Date") + ylab("Frass P Concentration (%)")
    plot(p)

```

Frass production has a great time-series record of P concentration data available. We may be able to compute its CO2 by time trend and check its statistical significance. 



### 3.7 Microbial P concentration
``` {r, echo=FALSE, message=FALSE, warning=FALSE}
    ### calculate the variable
    microbial_p_concentration <- make_microbial_p_concentration()
    
    ### assign CO2 treatment labels
    for (i in aCO2_rings) {
        microbial_p_concentration[microbial_p_concentration$Ring == i, "CO2"] <- "aCO2"
    }
    for (i in eCO2_rings) {
        microbial_p_concentration[microbial_p_concentration$Ring == i, "CO2"] <- "eCO2"
    }
    
    ### make the ring over date plot
    p <- ggplot(microbial_p_concentration,
                aes(Date, PercP, color=factor(CO2))) +   
            geom_point(size = 5) +
            xlab("Date") + ylab("Microbial P Concentration (%)")
    plot(p)

```

Note, this is based on 0-10cm data only, as this is the only depth with data available. It's not consistent with any other soil related variables (0-30cm). 

### 3.8 Soil P concentration
``` {r, echo=FALSE, message=FALSE, warning=FALSE}
    ### calculate the variable
    soil_p_concentration <- make_soil_p_concentration()
    
    ### assign CO2 treatment labels
    for (i in aCO2_rings) {
        soil_p_concentration[soil_p_concentration$Ring == i, "CO2"] <- "aCO2"
    }
    for (i in eCO2_rings) {
        soil_p_concentration[soil_p_concentration$Ring == i, "CO2"] <- "eCO2"
    }
    
    ### make the ring over date plot
    p <- ggplot(soil_p_concentration,
                aes(Date, PercP, color=factor(CO2))) +   
            geom_point(size = 5) +
            xlab("Date") + ylab("Soil P Concentration (%)")
    plot(p)

```

Note, soil data are available at various depths (0-10, 10-20 and 20-30cm), and data were not collected for all depths at different time points. 

``` {r, echo=FALSE} 

    soil_c_pool <- make_soil_c_pool(soil_bulk_density)
    ### make the depth over date plot
    p <- ggplot(soil_c_pool,
                aes(Date, TotC, color=factor(Depth))) +   
            geom_point(size = 5) +
            xlab("Date") + ylab("Soil C Pool at various depth")
    plot(p)
```

It is observed from the above C pool at various depths that, only two dates have C pool measured over three depths. We need to come up with a way to either extrapolate based on depth profile or to remove dates without 3-depth measurements. 

### 3.9 Soil Phosphate P concentration
Averaged across multiple depths, without consideration of assigning relative weight score. Need to update this calculation later. 

Also, soil phosphate P is the PO4-P, not PO4 concentration. 

Further, phosphate measurements were taken at quarterly timestep, but it should not reflect the accumulation of phosphate production over time time frame (I think). But this needs to be confirmed by soil people, as I am not sure whether the measurements excluded plants and mycorrhizaes. Even with exclusion, we still can't be sure this concentration reflects the total soil phosphate production over time period, as rainfall and hence leaching may also affect the concentration. 

``` {r, echo=FALSE, message=FALSE, warning=FALSE}
    ### calculate the variable
    soil_phosphate_concentration <- make_soil_phosphate_concentration()
    
    ### assign CO2 treatment labels
    for (i in aCO2_rings) {
        soil_phosphate_concentration[soil_phosphate_concentration$Ring == i, "CO2"] <- "aCO2"
    }
    for (i in eCO2_rings) {
        soil_phosphate_concentration[soil_phosphate_concentration$Ring == i, "CO2"] <- "eCO2"
    }
    
    ### make the ring over date plot
    p <- ggplot(soil_phosphate_concentration,
                aes(Date, PercP, color=factor(CO2))) +   
            geom_point(size = 5) +
            xlab("Date") + ylab("Soil P Concentration (%)")
    plot(p)

```

Note: Soil PO4 data is only available for top 0-10 cm of the soil!

What happens at that time point after 2015? Is it a leaching-induced event?


### 3.10 Mycorrhizal P concentration
We currently don't have mycorrhizal data available yet. 

# 4. Important variables. 

### 4.1 Soil bulk density across depths
``` {r, echo=FALSE, warning=FALSE}

    soil_bulk_density <- make_soil_bulk_density()

    ### make the plot
    p <- ggplot(soil_bulk_density,
                aes(ring, bulk_density_kg_m3, color=factor(Depth))) +   
            geom_point(size = 5) +
            xlab("Ring") + ylab("Bulk density (kg m-3)")
    plot(p)


```

### 4.2 Soil P mineralization rate
``` {r, echo=FALSE, warning=FALSE}

    soil_p_mineralization <- make_soil_p_mineralization_flux(soil_bulk_density)

    ### make the plot
    # compare eCO2 and aCO2 treatment
    soil_p_mineralization[soil_p_mineralization$ring == 1, "CO2"] <- "eCO2"
    soil_p_mineralization[soil_p_mineralization$ring == 4, "CO2"] <- "eCO2"
    soil_p_mineralization[soil_p_mineralization$ring == 5, "CO2"] <- "eCO2"
    
    soil_p_mineralization[soil_p_mineralization$ring == 2, "CO2"] <- "aCO2"
    soil_p_mineralization[soil_p_mineralization$ring == 3, "CO2"] <- "aCO2"
    soil_p_mineralization[soil_p_mineralization$ring == 6, "CO2"] <- "aCO2"
    
    
    p <- ggplot(soil_p_mineralization, aes(date, p_mineralization_mg_m2_d, color=factor(CO2))) +   
        geom_point(size = 5) +
        xlab("Date") + ylab("P mineralization rate (mg m-2 d-1)") + 
        ggtitle("P mineralization rate over time") 
    plot(p)


```

### 4.3 Leaf P retranslocation coefficient
We need to think about data collected from the same time point. 


### 4.4 C fraction for frass production
``` {r, echo=FALSE, warning=FALSE}

    frass_c_fraction <- make_frass_c_fraction()


    ### make the plot
    p <- ggplot(frass_c_fraction,
                aes(Date, frass_c_fraction, color=factor(Ring))) +   
            geom_point(size = 5) + geom_line() + 
            xlab("Date") + ylab("Fraction of C in Frass (%)") + 
            scale_color_manual(values=c("#FF7F50", "#00FFFF", "#6495ED",
                                        "#FF4040", "#8B0000", "#0000FF"))
    plot(p)


```

### 4.5 Aboveground leaf biomass, litterfall and frass production
``` {r, echo = FALSE, warning=FALSE, message=FALSE}
   # hidden variables
    #### Canopy related variables (SLA, LAI, Canopy biomass)
    lai_variable <- make_lai_variable()
    sla_variable <- make_sla_variable()
    canopy_biomass_pool <- make_canopy_biomass_pool(lai_variable, sla_variable)
    
    ### make the plot
    p <- ggplot(canopy_biomass_pool,
                aes(Date, leaf_pool, color=factor(Ring))) +   
            geom_point(size = 0.5) + geom_line() + 
            xlab("Date") + ylab("Canopy biomass (g m-2)") + 
            scale_color_manual(values=c("#FF7F50", "#00FFFF", "#6495ED",
                                        "#FF4040", "#8B0000", "#0000FF"))
    plot(p)
    

```

The above figure is a time-series canopy biomass plot, derived based on SLA and LAI measurements. 

Assuming C fraction of 0.5, we can derive the canopy C pool, as:
``` {r, echo=FALSE}
    
    ### make the plot
    p <- ggplot(canopy_biomass_pool,
                aes(Date, leaf_pool*0.5, color=factor(Ring))) +   
            geom_point(size = 0.5) + geom_line() + 
            xlab("Date") + ylab("Canopy C pool (g C m-2)") + 
            scale_color_manual(values=c("#FF7F50", "#00FFFF", "#6495ED",
                                        "#FF4040", "#8B0000", "#0000FF"))
    plot(p)

```

Next, I calculate litter fall fluxes:
``` {r, echo=FALSE, warning=FALSE, message=FALSE}
    #### Litter production (leaf, twig, bark, seed)
    litter_c_production_flux <- make_litter_c_flux(c_fraction)
    
    plotDF <- melt(litter_c_production_flux, id.vars = 1:2, measure= 3:6)
    
    ### make the plot
    p <- ggplot(plotDF,
                aes(Date, value, color=factor(variable))) +   
            geom_point(size = 0.5, aes(color=factor(Ring))) + geom_smooth() + 
            xlab("Date") + ylab("Litterfall (mg C m-2 d-1)") 
    plot(p)
    
```

Finally, frass production flux is:
``` {r, echo=FALSE, warning=FALSE, message=FALSE}
 
    #### Frass production
    frass_c_production_flux <- make_frass_c_production_flux()

    ### make the plot
    p <- ggplot(frass_c_production_flux,
                aes(Date, frass_production_flux, color=factor(Ring))) +   
            geom_point(size = 0.5) + geom_line() + 
            xlab("Date") + ylab("Frass C production (mg C m-2 d-1)") + 
            scale_color_manual(values=c("#FF7F50", "#00FFFF", "#6495ED",
                                        "#FF4040", "#8B0000", "#0000FF"))
    plot(p)
    
```


### 4.6 Fineroot biomass and production
``` {r, echo=FALSE, warning=FALSE, message=FALSE}
    #### Fineroot pools and production
    # this is total fineroot biomass for 0-30cm
    # currently implemented whole plant C fraction
    # need to update with fineroot specific C fraction value
    fineroot_c_pool <- make_fineroot_c_pool(c_fraction_fr)
    fineroot_c_production_flux <- make_fineroot_c_production_flux(c_fraction_fr)
    
    ### make the plot
    p <- ggplot(fineroot_c_pool,
                aes(Date, fineroot_pool, color=factor(Ring))) +   
            geom_point(size = 0.5) + geom_line() + 
            xlab("Date") + ylab("Fineroot C pool (g C m-2)") + 
            scale_color_manual(values=c("#FF7F50", "#00FFFF", "#6495ED",
                                        "#FF4040", "#8B0000", "#0000FF"))
    plot(p)
    
```

Fineroot production flux is:
``` {r, echo=FALSE, warning=FALSE, message=FALSE}

    ### make the plot
    p <- ggplot(fineroot_c_production_flux,
                aes(Date, fineroot_production_flux, color=factor(Ring))) +   
            geom_point(size = 0.5) + geom_line() + 
            xlab("Date") + ylab("Fineroot C production (mg C m-2 d-1)") + 
            scale_color_manual(values=c("#FF7F50", "#00FFFF", "#6495ED",
                                        "#FF4040", "#8B0000", "#0000FF"))
    plot(p)
    
```


### 4.7 Understorey C pool
``` {r, echo=FALSE, warning=FALSE, message=FALSE}
    
    #### Understorey aboveground biomass 
    understorey_c_pool <- make_understorey_aboveground_c_pool(c_fraction_ud,
                                                              strip_area)
    

```

# 5. P pools and fluxes

### 5.1 Soil P pool
``` {r, echo=FALSE, warning=FALSE}

    soil_p_pool <- make_soil_p_pool(p_conc=soil_p_concentration,
                                    bk_density=soil_bulk_density)

    ### make the plot
    p <- ggplot(soil_p_pool,
                aes(Date, soil_p_g_m2, color=factor(Ring))) +   
            geom_point(size = 5) + geom_line() +
            xlab("Date") + ylab("Soil P pool (g m-2)") + 
            scale_color_manual(values=c("#FF7F50", "#00FFFF", "#6495ED",
                                        "#FF4040", "#8B0000", "#0000FF"))
    plot(p)


```

### 5.2 Soil PO4-P pool

Note, this pool is more readily available to plants as the microbial p pool needs one extra step when extracting.
``` {r, echo=FALSE, warning=FALSE}

    soil_phosphate_pool <- make_soil_phosphate_pool(p_conc=soil_phosphate_concentration,
                                                    bk_density=soil_bulk_density)

    ### make the plot
    p <- ggplot(soil_phosphate_pool,
                aes(Date, soil_phosphate_p_g_m2, color=factor(Ring))) +   
            geom_point(size = 5) + geom_line() + 
            xlab("Date") + ylab("Soil PO4-P production (g m-2)") + 
            scale_color_manual(values=c("#FF7F50", "#00FFFF", "#6495ED",
                                        "#FF4040", "#8B0000", "#0000FF"))
    plot(p)


```

### 5.3 Microbial P pool
Note, this is for top 10 cm of the soil. 

``` {r, echo=FALSE, warning=FALSE}

    microbial_p_pool <- make_microbial_p_pool(p_conc=microbial_p_concentration,
                                              bk_density=soil_bulk_density)

    ### make the plot
    p <- ggplot(microbial_p_pool,
                aes(Date, microbial_p_g_m2, color=factor(Ring))) +   
            geom_point(size = 5) + geom_line() + 
            xlab("Date") + ylab("Soil PO4-P production (g m-2)")
    plot(p)


```

### 5.4 Canopy P pool
``` {r, echo=FALSE, warning=FALSE}

    canopy_p_pool <- make_canopy_p_pool(p_conc=canopy_p_concentration,
                                        biom=canopy_biomass_pool)

    ### make the plot
    p <- ggplot(canopy_p_pool,
                aes(Date, leaf_p_pool, color=factor(Ring))) +   
            geom_point(size = 5) + geom_line() + 
            xlab("Date") + ylab("Canopy P (g m-2)") + 
            scale_color_manual(values=c("#FF7F50", "#00FFFF", "#6495ED",
                                        "#FF4040", "#8B0000", "#0000FF"))
    plot(p)


```

### 5.5 Wood P pool

``` {r, echo=FALSE, warning=FALSE}

    wood_p_pool <- make_wood_p_pool(p_conc=wood_p_concentration,
                                    c_pool=wood_c_pool)

    ### make the plot
    p <- ggplot(wood_p_pool,
                aes(Date, wood_p_pool, color=factor(Ring))) +   
            geom_point(size = 5) + geom_line() + 
            xlab("Date") + ylab("Wood P (g m-2)") + 
            scale_color_manual(values=c("#FF7F50", "#00FFFF", "#6495ED",
                                        "#FF4040", "#8B0000", "#0000FF"))
    plot(p)


```

We have only concentration measurement for one point in time. Assume this rate applies to all biomass. 
Check, why inconsistent time points. 

### 5.6 Fine root P pool
``` {r, echo=FALSE, warning=FALSE}

    fineroot_p_pool <- make_fineroot_p_pool(p_conc=fineroot_p_concentration,
                                            c_pool=fineroot_c_pool)

    ### make the plot
    p <- ggplot(fineroot_p_pool,
                aes(Date, fineroot_p_pool, color=factor(Ring))) +   
            geom_point(size = 5) + geom_line() + 
            xlab("Date") + ylab("Fineroot P (g m-2)") + 
            scale_color_manual(values=c("#FF7F50", "#00FFFF", "#6495ED",
                                        "#FF4040", "#8B0000", "#0000FF"))
    plot(p)


```


### 5.7 Understorey aboveground P pool
Still need to process Mathias's data.


### 5.8 Mycorrhizal P pool
Data still not available. 


### 5.9 Canopy P flux
Need to take into consideration leaf turnover. 

### 5.10 Wood P flux
Data incomplete. 

### 5.11 Fineroot production P flux
``` {r, echo=FALSE, warning=FALSE}

    fineroot_p_production <- make_fineroot_p_production(p_conc=fineroot_p_concentration,
                                                        c_flux=fineroot_c_production_flux)
    ### make the plot
    p <- ggplot(fineroot_p_production,
                aes(Date, fineroot_p_flux_mg_m2_d, color=factor(Ring))) +   
            geom_point(size = 5) + geom_line() + 
            xlab("Date") + ylab("Fineroot P production (,g m-2 d-1)") + 
            scale_color_manual(values=c("#FF7F50", "#00FFFF", "#6495ED",
                                        "#FF4040", "#8B0000", "#0000FF"))
    plot(p)


```

### 5.12 Leaflitter P flux
``` {r, echo=FALSE, warning=FALSE, message = FALSE}

    leaflitter_p_flux <- make_leaflitter_p_flux(p_conc=leaflitter_p_concentration)  

    ### make the plot
    p <- ggplot(leaflitter_p_flux,
                aes(Date, leaflitter_p_flux_mg_m2_d, color=factor(Ring))) +   
            geom_point(size = 5) + geom_line() + 
            xlab("Date") + ylab("Leaf litter P (mg m-2 d-1)") + 
            scale_color_manual(values=c("#FF7F50", "#00FFFF", "#6495ED",
                                        "#FF4040", "#8B0000", "#0000FF"))
    plot(p) 


```

### 5.13 other aboveground litter P flux
``` {r, echo=FALSE, warning=FALSE, message = FALSE}

    other_litter_p_flux <- make_other_litter_p_flux(p_conc=wood_p_concentration)  

    ### make the plot
    p <- ggplot(other_litter_p_flux,
                aes(Date, other_litter_p_flux_mg_m2_d, color=factor(Ring))) +   
            geom_point(size = 5) + geom_line() + 
            xlab("Date") + ylab("Other litter P (mg m-2 d-1)") + 
            scale_color_manual(values=c("#FF7F50", "#00FFFF", "#6495ED",
                                        "#FF4040", "#8B0000", "#0000FF"))
    plot(p)


```

### 5.14 Frass production P flux
``` {r, echo=FALSE, warning=FALSE}

    frass_p_production <- make_frass_p_production_flux(p_conc=frass_p_concentration,
                                                   c_flux=frass_c_production_flux,
                                                   c_frac=frass_c_fraction)

    ### make the plot
    p <- ggplot(frass_p_production,
                aes(Date, frass_p_flux_mg_m2_d, color=factor(Ring))) +   
            geom_point(size = 5) + geom_line() + 
            xlab("Date") + ylab("Frass P production (mg m-2 d-1)") + 
            scale_color_manual(values=c("#FF7F50", "#00FFFF", "#6495ED",
                                        "#FF4040", "#8B0000", "#0000FF"))
    plot(p)


```

### 5.15 Understorey production P flux
Needs more understorey biomass data to proceed. 

### 5.16 Soil mineralization P flux
``` {r, echo=FALSE, warning=FALSE}

    soil_p_mineralization <- make_soil_p_mineralization_flux(soil_bulk_density)

    ### make the plot
    p <- ggplot(soil_p_mineralization,
                aes(date, p_mineralization_mg_m2_d, color=factor(ring))) +   
            geom_point(size = 5) + 
            xlab("Date") + ylab("Soil P mineralization (mg m-2 d-1)") + 
            scale_color_manual(values=c("#FF7F50", "#00FFFF", "#6495ED",
                                        "#FF4040", "#8B0000", "#0000FF"))
    plot(p)


```

Positive values indicate minerlization, negative indicate immobilization. 

# 6. Summary tables
``` {r, echo=FALSE, message=FALSE, warning=FALSE}
    library(knitr)

```
Below, I summarize the per ring, per treatment table for all P concentration variables. 
Rings 2,3,6 are ambient CO2 rings, and rings 1,4,5 are elevated CO2 rings. 

``` {r, echo=FALSE} 
    source("programs/make_conc_summary_table_by_treatment.R")
    summary_table_concentration_by_treatment <- make_conc_summary_table_by_treatment()
    
    # make the table looks nicer
    colnames(summary_table_concentration_by_treatment)[1] <- "Variable"
    summary_table_concentration_by_treatment[,2:7] <- round(summary_table_concentration_by_treatment[,2:7], 3)
    
    kable(summary_table_concentration_by_treatment)

```

<br/>
<br/>
<br/>
<br/>

Next, I summarize all pool variables:
``` {r, echo = FALSE} 
    source("programs/make_pool_summary_table_by_treatment.R")
    summary_table_pool_by_treatment <- make_pool_summary_table_by_treatment()
    
    # make the table looks nicer
    colnames(summary_table_pool_by_treatment)[1] <- "Variable"
    summary_table_pool_by_treatment[,2:7] <- round(summary_table_pool_by_treatment[,2:7], 3)
    
    kable(summary_table_pool_by_treatment)

```

<br/>
<br/>
<br/>
<br/>

Finally, I summarize all flux variables:
``` {r, echo = FALSE} 
    source("programs/make_flux_summary_table_by_treatment.R")
    summary_table_flux_by_treatment <- make_flux_summary_table_by_treatment()
    
    # make the table looks nicer
    colnames(summary_table_flux_by_treatment)[1] <- "Variable"
    summary_table_flux_by_treatment[,2:7] <- round(summary_table_flux_by_treatment[,2:7], 3)
    
    kable(summary_table_flux_by_treatment)

```

<br/>
<br/>
<br/>
<br/>


# 7. C:P ratios

<br/>
<br/>
<br/>
<br/>
<br/>
<br/>


# 8. To-do list
- Continuous understorey C data (Matthias)
- Hedley Fractionation (Cat)
- Microbial turnover rate (Cat and Johanna, but may not be available soon)
- Mycorrhizal data (Jeff)
- Fine root mortality and turnover (Juan)
- Canopy turnover 
- Meteorological data, including soil moisture (HIEv)
- Summarize P concentration using min and max
- Calculate P budgeting variables and interpret the results. 
- Possibly look at 2015 to see CO2 treatment effect. 
# The end.
