---
title: "EucFACE P Synthesis"
author: "Mingkai Jiang"
date: "1/24/2018"
output: html_document
---


# 1. Description

This is the progress report of the EucFACE phosphorus (P) synthesis project. All results presented in this report is automatically linked with the code and the data. 

This file attempts to summarize P-related pools, fluxes, concentrations, their treatment and inter-ring variability, and their temporal patterns. 

The objective of this project is to have a proper account of the P balance for the EucFACE experiment, and to investigate the P by CO2 interactions over time (if enough data is available).

All the data used in this study are available at HIEv (at least that's the goal). 


# 2. Methods
### 2.1 Procedure
This procedure is a snapshot of the work. It might be easier just to follow the codes. 

    - Create a list of variables for the P synthesis.
    - For each variable, only consider ring, date, and sometimes depth as independent factors.
    - Calculate P concentrations first.
    - With Carbon balance project, calculate P pools and fluxes.
    - Calculate P budgeting variables, e.g. total standing P stock, P retranslocation coefficient, etc.
    - Make summary tables.
    - Interpret data, e.g. time-series trend, treatment effect, extreme response point, etc. 

### 2.2 Variables
Here, I am summarizing all the P-related pool, flux and concentration variables. 
Currently, this list is in-comprehensive. Will update continuously as the project proceeds. 

    - Soil P concentration and pool (different chemical forms and accessibilities)
    - Microbial P concentration and pool
    - Mycorrhizal P concentration and pool
    - Canopy P concentration, production and pool
    - Wood P concentration, production and pool
    - Fineroot P concentration, production and pool
    - Litterfall P concentration and production
    - P mineralization rate
    - Frass P concentration and production
    - Understorey P concentration, production and pool
    - Standing P stock
    - P requirement
    - Total P retranslocation 
    - Leaf P retranslocation coefficient
    - P uptake
    - P uptake / requirement
    - MRT
    - Standing PUE
    

# 3. P concentration calculations. 
```{r, echo=FALSE, message=FALSE, warning=FALSE}
### source the functions to start the calculations.
source("programs/prepare.R")

### assign aco2 and eco2 rings
aCO2_rings <- c(2,3,6)
eCO2_rings <- c(1,4,5)

```

### 3.1 Canopy P concentration
Canopy P concentration summarized by ring over the experimental period:
``` {r, echo=FALSE}
### calculate the variable
canopy_p_concentration <- make_canopy_p_concentration()

### make the ring over date plot
p <- ggplot(canopy_p_concentration,
            aes(Date, PercP, color=factor(Ring))) +   
        geom_point(size = 5) +
        xlab("Date") + ylab("Canopy P Concentration (%)")
plot(p)

```

Canopy P concentration summarized by CO2 treatment over the experimental period:
``` {r, echo=FALSE}
### assign CO2 treatment labels
for (i in aCO2_rings) {
    canopy_p_concentration[canopy_p_concentration$Ring == i, "CO2"] <- "aCO2"
}
for (i in eCO2_rings) {
    canopy_p_concentration[canopy_p_concentration$Ring == i, "CO2"] <- "eCO2"
}

### make the treatment over date plot
p <- ggplot(canopy_p_concentration,
            aes(Date, PercP, color=factor(CO2))) +   
        geom_point(size = 5) +
        xlab("Date") + ylab("Canopy P Concentration (%)")
plot(p)

```

Visually, there is **no** CO2 treatment effect. 



### 3.2 Wood P concentration
Wood P concentration summarized by treatment, and we only have one date:

``` {r, echo=FALSE}
### calculate the variable
wood_p_concentration <- make_wood_p_concentration()

### assign CO2 treatment labels
for (i in aCO2_rings) {
    wood_p_concentration[wood_p_concentration$Ring == i, "CO2"] <- "aCO2"
}
for (i in eCO2_rings) {
    wood_p_concentration[wood_p_concentration$Ring == i, "CO2"] <- "eCO2"
}

### make the ring over date plot
p <- ggplot(wood_p_concentration,
            aes(CO2, PercP)) +   
        geom_point(size = 5) +
        xlab("Treatment") + ylab("Wood P Concentration (%)")
plot(p)

```

Visually, we can't distinguish the difference between CO2 treatment.



### 3.3 Fine root P concentration

There are multiple depths (i.e. 0-10cm, 10-30cm), currently we are averaging the concentrations over depths. 

Fineroot P concentration summarized by treatment over time:

``` {r, echo=FALSE}
### calculate the variable
fineroot_p_concentration <- make_fineroot_p_concentration()

### assign CO2 treatment labels
for (i in aCO2_rings) {
    fineroot_p_concentration[fineroot_p_concentration$Ring == i, "CO2"] <- "aCO2"
}
for (i in eCO2_rings) {
    fineroot_p_concentration[fineroot_p_concentration$Ring == i, "CO2"] <- "eCO2"
}

### make the treatment over date plot
p <- ggplot(fineroot_p_concentration,
            aes(Date, PercP, color=factor(CO2))) +   
        geom_point(size = 5) +
        xlab("Date") + ylab("Fineroot P Concentration (%)")
plot(p)

```


### 3.4 Leaflitter P concentration

### 3.5 Understorey P concentration

### 3.6 Frass P concentration

### 3.7 Microbial P concentration

### 3.8 Soil P concentration

### 3.9 Soil Phosphate P concentration

### 3.10 Mycorrhizal P concentration



```{r, echo=FALSE}

```



# The end
